Improved Approximation Algorithms for the Maximum Carpool Matching Problem

### 2017-01-31 src/intro.tex
fix subcaption, no warnings

Given a directed graph $G$ and a subset $M \subseteq A$, define
$\din[M](v) \eqdf \abs{\set{u :(u,v) \in M}}$ and
$\dout[M](v) \eqdf \abs{\set{u :(v,u) \in M}}$.
%
A feasible \emph{carpool matching} is a subset of arcs, $M \subseteq
A$, such that every $v \in V$ satisfies:%
\begin{inparaenum}[(i)]
\item $\din[M](v) \cdot \dout[M](v) = 0$,
\item $\din[M](v) \leq c(v)$, and 
\item $\dout[M](v) \leq 1$.
\end{inparaenum}
A vertex $v$ for which $\dout[M](v) = 1$ is a \emph{passenger}, and a
vertex for which $\dout[M](v) = 0$ is a driver who has $\din[M](v)$
passengers.  In the \carpool problem the goal is to find a matching
$M$ of maximum total weight, namely to minimize
$w(M) \eqdf \sum_{(v,u) \in M} w(v,u)$.  In other words, the \carpool
problem is about finding a set of (directed toward the center) vertex
disjoint stars that maximizes the total weight of the arcs.
Figure~\ref{fig:carpool} is an example of the \carpool problem.

\begin{figure}
\centering
\input{fig-carpool}
\caption[]{
\label{fig:carpool}
A carpool matching example: 
\subref{subfloat:input} 
a directed graph with capacities on the vertices and weights on the arcs. 
\subref{subfloat:output}
a feasible matching with total weight of 26.
$P$ is the set of blue vertices, and $D$ is the set of red, dashed vertices. 
}
\end{figure}

In this paper we also consider an extension of \carpool, called \gcp,
in which each vertex represents a group of passengers.  Group sizes
are non-uniform, and we consider the case where a group size may
change as a function of the driver.  Formally, each arc $(u,v) \in A$
is associated with a size $s(u,v) \in \N$, and the constraint
$\din[M](v) \leq c(v)$ is replaced with the constraint
$\sum_{u:(u,v) \in M} s(u,v) \leq c(v)$.
%
Notice that \textsc{Knapsack} is the special case, where only arcs
directed at a single vertex have non-zero weights.

@@

\subparagraph{Related work.}
%
Hartman et al.~\cite{hartman2013optimal} proved that the \carpool
problem considered in this paper is NP-hard, and that the problem
remains NP-hard even for a binary weight function when the capacity
function $c(v) \leq 2$ for every vertex in $V$.  It is also worth
mentioning, that in the undirected, uncapacitated, unweighted variant
of the problem, the set of drivers in an optimal solution form a
minimum dominating set (see Section~\ref{sec:hardness}).  When the set
of drivers is known in advanced, however, the problem becomes
tractable and can be solved using a reduction to a flow network
problem.

@@

Kutiel~\cite{Kutiel16}

\todo[inline]{add a paragraph about the paper from arXiv}

\todo[inline]{Mention Goerge's paper with Amotz (ESA)}

@@

\subparagraph{Our contribution.}
%
In Section~\ref{sec:sub} we show that the problem can be formulated as
unconstrained submodular maximization problem, thus it has a
randomized $2$-approximation algorithm due to~\cite{BFNS15}.  We also
provide a deterministic $(2+\eps)$-approximation algorithm, for any
$\eps>0$, which is based on local search.

In Section~\ref{sec:local} we consider the unweighted variant of the
problem.  We provide a $(\frac{2\cmax}{\cmax+1} + \eps)$-approximation
algorithm, for any $\eps>0$, where $\cmax$ is the maximum capacity.
We also show that \carpool is APX-hard even if $\cmax \leq 3$.
%
In Section~\ref{sec:group} we consider \gcp.  We first show that \gcp
cannot be formulated as an unonstrained submodular maximization
problem as \carpool.  We show, however, that this problem still admits
a $(2 + \eps)$-approximation algorithm, and argue that the local
search algorithm generalizes to \gcp as well.

>>>>

Given a directed graph $G$ and a subset $M \subseteq A$, define
$\din[M](v) \eqdf \abs{\set{u :(u,v) \in M}}$ and
$\dout[M](v) \eqdf \abs{\set{u :(v,u) \in M}}$.
%
A feasible \emph{carpool matching} is a subset of arcs, $M \subseteq
A$, such that every $v \in V$ satisfies:
\begin{inparaenum}[(i)]
\item $\din[M](v) \cdot \dout[M](v) = 0$,
\item $\din[M](v) \leq c(v)$, and 
\item $\dout[M](v) \leq 1$.
\end{inparaenum}
A vertex $v$ for which $\dout[M](v) = 1$ is a \emph{passenger}, and a
vertex for which $\dout[M](v) = 0$ is a driver who has $\din[M](v)$
passengers.  In the \carpool problem the goal is to find a matching
$M$ of maximum total weight, namely to maximize
$w(M) \eqdf \sum_{(v,u) \in M} w(v,u)$.  In other words, the \carpool
problem is about finding a set of (directed toward the center) vertex
disjoint stars that maximizes the total weight of the arcs.
Figure~\ref{fig:carpool} is an example of the \carpool problem.

\begin{figure}
\centering
\input{fig-carpool}
\caption[]{
\label{fig:carpool}
A carpool matching example: 
\subref{carpool:input}
a directed graph with capacities on the vertices and weights on the arcs. 
\subref{carpool:output}
a feasible matching with total weight of 26.
$P$ is the set of blue vertices, and $D$ is the set of red, dashed vertices. 
}
\end{figure}

In this paper we also consider an extension of \carpool, called \gcp,
in which each vertex represents a group of passengers, and  
each group might have a different size. 
Formally, each vertex $u \in V$ has a size $s(u) \in \N$, 
and the constraint $\din[M](v) \leq c(v)$ is replaced with the constraint
$\sum_{u:(u,v) \in M} s(u) \leq c(v)$.
%
Notice that \textsc{Knapsack} is the special case, where only
arcs directed at a single vertex have non-zero (integral) weights.

@@

\subparagraph{Related work.}
%
Hartman et al.~\cite{hartman2013optimal} proved that the \carpool
problem is NP-hard, and that the problem
remains NP-hard even for a binary weight function when the capacity
function $c(v) \leq 2$ for every vertex in $V$.  It is also worth
mentioning, that in the undirected, uncapacitated, unweighted variant
of the problem, the set of drivers in an optimal solution form a
minimum dominating set (see Section~\ref{sec:hardness}).  When the set
of drivers is known in advanced, however, the problem becomes
tractable and can be solved using a reduction to a flow network
problem.

@@

Kutiel~\cite{kutiel2016} presented a $\frac{1}{3}$-approximation
algorithm for the problem and $\frac{1}{2}$-approximation algorithm for the
unweighted variant of the problem.

Bar-Noy et al.~\cite{bar2015improved} considered the problem of partition a
graph into disjoint paths of length 2.
This problem is a special case of the undirected carpool matching
where all capacities are exactly 2.
They show a $.5833$, and $\frac{3}{4}$-approximation algorithms for the weighted
and unweighted variants of the problem respectively.

For the general set packing problem, the best known approximation algorithm is
due to Chandra et al.~\cite{chandra2001greedy} and it admits a 
$\frac{2(k +1)}{3}$ approximation ratio where $k$ is the size of the sets.

@@

\subparagraph{Our contribution.}
%
In Section~\ref{sec:sub} we show that the \carpool can be formulated as
unconstrained submodular maximization problem, thus it has a
$\frac{1}{2}$-approximation algorithm due to~\cite{BFNS15}.

In Section~\ref{sec:local} we consider the unweighted variant of the
problem.  We provide a $(\frac{2\cmax}{\cmax+1} + \eps)$-approximation
algorithm, for any $\eps>0$, where $\cmax$ is the maximum capacity.
We also show that \carpool is APX-hard even if $\cmax \leq 3$.
%
In Section~\ref{sec:group} we consider \gcp.  
We first show that the unconstrained submodular maximization formulation for
the \carpool does not follow for the \gcp.
We show, however, that this problem still admits
a $(\frac{1}{2} -\varepsilon)$-approximation algorithm, and argue that the local
search algorithm generalizes to \gcp as well.

### 2017-01-31 src/main.tex
fix subcaption, no warnings

\usepackage{amsmath,amssymb,mathtools,amsthm}
\usepackage[linesnumbered, ruled]{algorithm2e}
\usepackage{xspace}
\usepackage{subfig}
\usepackage{paralist}
%\usepackage{hyperref}
%\hypersetup{final}
%\usepackage[inline]{enumitem}

% DEBUG
%\usepackage{fullpage}
\usepackage{todonotes}
%\usepackage{lineno}
%\linenumbers

% EXTRA
%\usepackage{authblk}

@@

\title{Improved Approximation Algorithms for the
Maximum Carpool Matching Problem}

\titlerunning{Improved Approximation Algorithms for Maximum Carpool Matching}

\author[1]{Gilad Kutiel}
\author[2]{Dror Rawitz%
\thanks{Supported in part by the Israel Science Foundation (grant no.~497/14).}}

\affil[1]{Department of Computer Science, Technion, Haifa,
  Israel. \\ \texttt{gkutiel@cs.technion.ac.il}}
\affil[2]{Faculty of Engineering, Bar Ilan University, Ramat Gan,
  Israel. \\ \texttt{dror.rawitz@biu.ac.il}}

\authorrunning{G. Kutiel and D. Rawitz}

\Copyright{G. Kutiel and D. Rawitz}

\subjclass{F.2.2 Nonnumerical Algorithms and Problems}

\keywords{approximation algorithms, local search}

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Acces}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CVIT 2016}
\EventAcronym{CVIT}
\EventYear{2016}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
% Editor-only macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%\bibliographystyle{plain}
\bibliographystyle{plainurl}% the recommended bibstyle
\bibliography{main}

>>>>

\usepackage{amsmath,amssymb,mathtools,amsthm}
\usepackage[linesnumbered, ruled]{algorithm2e}
\usepackage{xspace}
\usepackage{subcaption}
\usepackage{paralist}

% DEBUG
\usepackage{todonotes}

@@

\input{in/title}
%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Acces}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CVIT 2016}
\EventAcronym{CVIT}
\EventYear{2016}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
% Editor-only macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\bibliographystyle{plainurl}
\bibliography{main}
